## About Data

```{r}
source("01_mlr3-experiments.R")
```

Preprocessed `ds801_geochem1km.rds` regression matrix:

-   has all chemical sample variables (potential targets), and aligned covariates
-   "grouping" columns of interest: `site_id`, `ID`, `state`

```{r plot-ds801}
ds801$empty <- 0

plot(ds801[,"geometry"])

```


### Questions: 

-   What is the difference between `ID` and `site_id` in `ds801_geochem1km.rds`?
    -   maybe `ID` is the 30x30km blocks as mentioned in @eml-tutorial? since less unique IDs than `site_id`

```{r}
ds801 |> 
  dplyr::select(site_id, ID) |> 
  dplyr::summarise(across(c(site_id, ID), dplyr::n_distinct))
```

```{r}
plot(chicago[,"snow.occ_2000..2012_p90_aug_usa48"])
```


-   Why `chicago_grid1km.rds`?
    -   has basically the same columns as the RF regression matrix created in @eml-tutorial except the extra grouping variables, and geometry 
    - `x` and `y` are coordinates
    
```{r}
setdiff(names(ds801_pb), names(chicago))
setdiff(names(chicago), names(ds801_pb))
```
    
